---
title: "Visualizing and data wrangling with the `timetk` package for R"
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

```{r setup}
library(tidyverse)
library(tidymodels)
library(tidyquant)

# remotes::install_github("business-science/timetk")
library(timetk)
```

## Loading Stock Data with `tidyquant`

For our data, we'll create a portfolio of 5 ETFs: SPY, EFA, IJS, EEM, and AGG.

First, we'll load the data from Yahoo Finance using `tq_get()` and will calculate daily returns using `tq_transmute()`. 

```{r data}
# The symbols vector holds our tickers. 
symbols <- c("SPY","EFA", "IJS", "EEM","AGG")

etf_daily_returns <- 
  symbols %>% 
  tq_get(get = "stock.prices", from = "2004-01-01") %>% 
  group_by(symbol) %>% 
  tq_transmute(adjusted, periodReturn, period = "daily", col_rename = "returns")
```


Next, we'll build a portfolio with the following weights:

* SPY: 25%
* EFA 25%
* IJS 25%
* EEM 20%
* AGG 5%

```{r Aggressive Portfolio}
w_agg <- c(0.25, 0.25, 0.25, 0.20, 0.05)

portfolio_returns <- 
  etf_daily_returns %>% 
  tq_portfolio(assets_col = symbol, 
               returns_col = returns, 
               weights = w_agg,
               col_rename = "returns")
```



## Visualize Time Series Data

We can create a nice looking plot of our time series data with a single line of code.

```{r}
portfolio_returns %>%
  plot_time_series(date, returns, .color_var = year(date),
                   .interactive = FALSE, .color_lab = "Year")
```

We can even make this plot interactive using `timetk`'s integration with `plotly`:

```{r}
portfolio_returns %>%
  plot_time_series(date, returns, .color_var = year(date),
                   .interactive = TRUE, .plotly_slider = TRUE, .color_lab = "Year")
```


`timetk` will automatically facet plots with grouped data:

```{r}
etf_daily_returns %>% 
  group_by(symbol) %>% 
  plot_time_series(date, returns, .facet_ncol = 2, .color_var = symbol, .interactive = FALSE)
```


`timetk` also has a built-in function for [plotting anomaly diagnostics](https://business-science.github.io/timetk/reference/plot_anomaly_diagnostics.html):

```{r}
portfolio_returns %>% 
  plot_anomaly_diagnostics(date, returns)
```


To adjust outlier detection, change the `alpha` parameter.

Anomaly detection is done by the `tk_anomaly_diagnostics` function:

```{r}
portfolio_returns %>% 
  tk_anomaly_diagnostics(date, returns)
```


### Shiny with `timetk`

We can use these visualization tools to easily build a simple [**Shiny application**](https://nrohr.shinyapps.io/timetk-app/) that helps us diagnose time series data.


## Data Wrangling

### Summarizing and filtering by time

`timetk` also includes some functions that work like common `dplyr` verbs, but are especially useful when working with time series data.

Let's look at our individual ETFs to apply some of the data wrangling tools from `timetk`.

```{r}
etf_daily_returns %>% 
  group_by(symbol) %>% 
  plot_time_series(date, returns, .facet_ncol = 2, .interactive = FALSE)
```


We can use `summarise_by_time()` to aggregate over a specific time period. For example, let's smooth our returns by averaging them over each month:

```{r}
etf_daily_returns %>% 
  summarise_by_time(
    date, .by = "month",
    returns = AVERAGE(returns)
  ) %>% 
  plot_time_series(date, returns, .facet_ncol = 2, .interactive = FALSE)
```


We can also filter by time range using the `filter_by_time()` function:

```{r}
etf_daily_returns %>% 
  filter_by_time(date, "2012-06", "2020") %>% 
  summarise_by_time(
    date, .by = "month",
    returns = AVERAGE(returns)
  ) %>% 
  plot_time_series(date, returns, .facet_ncol = 2, .interactive = FALSE)
```


### Padding and imputing missing data

`timetk` also has functions for padding and imputing missing data.

Our tibbles of data are missing some dates - weekends and holidays. In our ETF returns:

```{r}
etf_daily_returns
```

Some packages for working with time series data require there to be no missing entries. Luckily, `timetk` makes it easy to add placeholders for any msising dates in our data.

The `pad_by_time()` function will replace missing values with `NA` by default.

```{r}
etf_daily_returns %>% 
  pad_by_time(date, .by = "auto")
```


In some cases, we may want to impute missing values (i.e., replace `NA` with an actual value). This also allows us to impute more granular data, like hourly stock prices.

Let's start with the daily closing prices for SPY:

```{r}
spy_prices <- 
  tq_get("SPY", get = "stock.prices", from = "2004-01-01") 

spy_prices

spy_prices %>% 
  plot_time_series(date, close, .interactive = FALSE)
```


Now let's impute hourly closing prices for every day of the year using `ts_impute_vec()`. (This is done linearly when `period = 1`)

```{r}
spy_prices_hourly <- 
  spy_prices %>% 
  pad_by_time(date, .by = "hour") %>% # pad with additional observations for every hour
  mutate_at(vars(close), .funs = ts_impute_vec, period = 1)

spy_prices_hourly
```


### Sliding / rolling calculations

Rolling functions are also useful for working with time series data. For example, we might want to calculate a rolling average of our daily portfolio returns. This can be done using the `slidify()` function:

```{r}
roll_avg <- slidify(.f = AVERAGE, .period = 7, .align = "center", .partial = TRUE)

portfolio_returns %>% 
  mutate(rolling_avg = roll_avg(returns)) %>%
  pivot_longer(cols = c(returns, rolling_avg)) %>% 
  plot_time_series(date, value, .color_var = name, .smooth = FALSE, .plotly_slider = TRUE)
```


